The Chemistry Development Kit (CDK): atom typing, rendering, molecular formulas, and substructure searching

### 2016-10-09 article.tex
More tweaks

Also, although fingerprints traditionally are bit vectors a count
fingerprint was also introduced making fingerprints based on integer
vectors supported in CDK as well. This is useful for fingerprints that are based
on substructures, such as the \texttt{SignatureFingerprinter} which uses the
aforementioned atomic signatures. The counts in the fingerprint then represent
how often this substructure is found in the molecule it represents. For
example, for signature depth 1:

@@

As the library grew over the years, so did the complexity of the
maintenance. The main branch frequently failed to compile and bug
fixes became more onerous due to unexpected side effects.  Often
fixing a bug in one part of the code, broke some other code which made
the incorrect assumptions about the fixed code. With the increase in
the CDK developer community, such issues were inevitable, in the
absence of any formal coding and testing standards.

To address these issues, we have adopted a number of coding
standards. While not a comprehensive implementation of software
engineering best practices, they attempt to find a balance between
increasing code maintainability and being flexible enough to allow
efficient code development. However, we appreciate the subjective
nature of this statement, and some adopted guidelines have been
heavily discussed and debated in the CDK community.

Arguably, perhaps the biggest factor in improved code quality is a
peer review process where any functionality changing patch is required
to be reviewed by one independent, senior CDK developer for the
development branch, and two reviewers for stable branches. This patch
development system is supported by a number of automated validations
steps as outlined below.

The next sections describe some approaches the project have adopted that allows
us to maintain the CDK library as it is today.

@@

One of the key approaches we have adopted, is to make the CDK more modular. The CDK assigns
every class to a module, and defines dependencies between modules. For example, core modules
are not allowed to depend on a module holding data classes implementing the CDK interfaces;
instead, they may only depend on the interfaces themselves. This, for example, is to ensure
that dependencies are minimized and to make it easier to exclude CDK functionality with
third-party dependencies that are not needed.

An overview of key modules with description, important changes, and dependencies
on third-party libraries is given in Table~4\ref{tab:modules} and the dependencies
between the CDK modules are depicted in
Figure~\ref{fig:deps}.\todo{the dependency graph is not very legible}

>>>>

Also, although fingerprints traditionally are bit vectors a count
fingerprint was also introduced making fingerprints based on integer
vectors supported in CDK as well. This is useful for fingerprints that are based
on substructures, such as the \texttt{SignatureFingerprinter} which uses the
aforementioned atomic signatures. The counts in the fingerprint then represent
how often this substructure is found in the molecule it represents. For
example, for a signature of depth 1:

@@

As the CDK library grew over the years, so did the complexity of the
maintenance. The main branch frequently failed to compile and bug
fixes became more onerous due to unexpected side effects.  Often
fixing a bug in one part of the code, broke some other code which made
the incorrect assumptions about the fixed code. With the increased size of
the CDK developer community, such issues were inevitable in the
absence of any formal coding and testing standards.

To address these issues, we have adopted a number of coding
standards. While not a comprehensive implementation of software
engineering best practices, they attempt to find a balance between
increasing code maintainability and being flexible enough to allow
efficient code development. We appreciate the subjective
nature of this statement, and some adopted guidelines have been
heavily discussed and debated in the CDK community.

Arguably, perhaps the biggest factor in improved code quality is a
peer review process where any functionality changing patch is required
to be reviewed by one independent, senior CDK developer for the
development branch, and by two reviewers for stable branches. This patch
development system is supported by a number of automated validations
steps as outlined below.
The next sections describe some approaches the project have adopted that allows
us to maintain the CDK library as it is today.

@@

One of the central approaches we have adopted, is to make the CDK more modular. The CDK assigns
every class to a module, and defines dependencies between modules. For example, core modules
are not allowed to depend on modules with data classes implementing the CDK interfaces;
instead, they may only depend on the interfaces themselves. This ensures that
dependencies are minimized. Furthermore, it also allows cherrypicking CDK
functionality, reducing the number of third-party library dependencies that are
needed.
An overview of key modules with description, important changes, and dependencies
on third-party libraries is given in Table~4\ref{tab:modules} and the dependencies
between the CDK modules are depicted in
Figure~\ref{fig:deps}.

% \todo{the dependency graph is not very legible}
% EW: not sure how to solve that
